---
title: "Lab 5"
author: 'Rich Chu, Lorena Mancino'
output: html_notebook
---

# Lab 5 Get Me out(lier) of Here!

## Problem 0

```{r}
x <- c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9) # x is the given data set 
```

## Problem 1

```{r}
quantile(x, 0.25) # calculates the first quartile of x
quantile(x, 0.75) # calculates the third quartile of x
```

## Problem 2

```{r}
#' Takes a vector containing a sample and returns a vector with three components: the first quartile, the third quartile, and the interquartile range.
#'
#' @param x a vector containing the data set 
#'
#' @return returns a list containing the first quartile, the third quartile, and the interquartile range.

quartiles <- function(x){
  first <- quantile(x, 0.25)
  third <- quantile(x, 0.75)
  list <- list(first, third, third - first)
  names(list) <- c('First', 'Third', 'Interquartile_Range')
  return(list)
}
quartiles(x)
```

## Problem 3

According to Tukey's rules, the lower bound for outliers is -18.6625 and there are no values less than that, meaning there are no lower outliers in the set.
According to Tukey's rules, the upper bound for outliers is 19.0375 and there is one value above than that, meaning that 28.7 is the only upper outliers in the set.

## Problem 4

```{r}
#' Takes a vector containing a sample and returns a Boolean vector with TRUE at the elements corresponding to outlier observations and FALSE everywhere else.
#' 
#' @param x a vector containing the data set 
#'
#' @return a vector of Boolean values 

tukey.outlier <- function(x){
  q <- quartiles(x)
  lower.bound <- q$First - (1.5 * q$Interquartile_Range)
  upper.bound <- q$Third + (1.5 * q$Interquartile_Range)
  
  bool <- vector(length=length(x))
  
  for(i in 1:length(x)){
    if(x[i] > upper.bound){
        bool[i] <- TRUE
      }
    else if(x[i] < lower.bound){
        bool[i] <- TRUE
    }
    else{
      bool[i] <- FALSE
    }
  }
  return(bool)
}

tukey.outlier(x)
```

## Problem 5

```{r}
#' Compares the outliers of a data set to its actual outliers 
#' 
#' @param FUN the function used to compute outliers with tukey.outliers as its default value 
#'
#' @return Returns TRUE if the function correctly identifies the outliers and FALSE otherwise.

test.tukey.outlier <- function(FUN= tukey.outlier){
  calculated.output <- tukey.outlier(x)
  
  if(all((calculated.output == FUN) == TRUE)){
    return(TRUE)    
  }
  else{
    return(FALSE)
  }
}
test.tukey.outlier(tukey.outlier(x))
```

## Problem 6

```{r}
x2 <- -x

tukey.outlier(x2)
```
The only data value that should be an outlier in -x is the eleventh term (-28.7). This makes sense because -x is just a reflection of x, so the outlier bounds are just flipped. 

## Problem 7

```{r}
#' Compares the outliers of a data set to its actual outliers 
#' 
#' @param FUN the function used to compute outliers with tukey.outliers as its default value 
#'
#' @return Returns a Boolean vector with two elements, one for x and one for -x 

test.tukey.outlier <- function(FUN= tukey.outlier){
  calculated.output <- tukey.outlier(x)
  negative.output <- tukey.outlier(-x) # new modification to the function from problem 5 
  return <- vector(length=2)
  
  if(all((calculated.output == FUN) == TRUE)){
    return[1]=TRUE    
  }
  else{
    return[1]=FALSE
  }
  
  if(all((negative.output == FUN) == TRUE)){
    return[2]=TRUE    
  }
  else{
    return[2]=FALSE
  }
  return(return)
}

```

## Problem 8

```{r}
test.tukey.outlier(tukey.outlier(x))
```

## Problem 9

```{r}
x3 <- 0.1*x

tukey.outlier(x3)
```
The only data value that should be an outlier in 0.1*x is the eleventh term (2.87). This makes sense because it's scaled down. 

## Problem 10

```{r}
#' Compares the outliers of a data set to its actual outliers 
#' 
#' @param FUN the function used to compute outliers with tukey.outliers as its default value 
#'
#' @return Returns a Boolean vector with three elements, one for x, one for -x, and one for 0.1*x 

test.tukey.outlier <- function(FUN= tukey.outlier){
  calculated.output <- tukey.outlier(x)
  negative.output <- tukey.outlier(-x)
  scaled.down.output <- tukey.outlier(0.1*x) # new modification to the function from problem 7 
  return <- vector(length=3)
  
  if(all((calculated.output == FUN) == TRUE)){
    return[1]=TRUE    
  }
  else{
    return[1]=FALSE
  }
  
  if(all((negative.output == FUN) == TRUE)){
    return[2]=TRUE    
  }
  else{
    return[2]=FALSE
  }
  
  if(all((scaled.down.output == FUN) == TRUE)){
    return[3]=TRUE    
  }
  else{
    return[3]=FALSE
  }
  
  return(return)
}
```

## Problem 11

```{r}
test.tukey.outlier(tukey.outlier(x))
```
















