---
title: "Lab 5"
author: 'Rich Chu, Lorena Mancino'
output: html_notebook
---

# Lab 5 Get Me out(lier) of Here!

## Problem 0

```{r}
x <- c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
```


## Problem 1

```{r}
quantile(x, 0.25)
quantile(x, 0.75)
```

## Problem 2

```{r}
quartiles <- function(x){
  first <- quantile(x, 0.25)
  third <- quantile(x, 0.75)
  list <- list(first, third, third - first)
  names(list) <- c('First', 'Third', 'Interquartile_Range')
  return(list)
}
quartiles(x)
```

## Problem 3
According to Tukey's rules, the lower bound for outliers is -18.6625 and there are no values less than that, meaning there are no lower outliers in the set.
According to Tukey's rules, the upper bound for outliers is 19.0375 and there is one value above than that, meaning that 28.7 is the only upper outliers in the set.

## Problem 4

```{r}
tukey.outlier <- function(x){
  q <- quartiles(x)
  lower.bound <- q$First - (1.5 * q$Interquartile_Range)
  upper.bound <- q$Third + (1.5 * q$Interquartile_Range)
  
  bool <- vector(length=length(x))
  
  for(i in 1:length(x)){
    if(x[i] > upper.bound){
        bool[i] <- TRUE
      }
    else if(x[i] < lower.bound){
        bool[i] <- TRUE
    }
    else{
      bool[i] <- FALSE
    }
  }
  return(bool)
}

tukey.outlier(x)
```

## Problem 5

```{r}
test.tukey.outlier <- function(FUN= tukey.outlier){
  calculated.output <- tukey.outlier(x)
  
  if(all((calculated.output == FUN) == TRUE)){
    return(TRUE)    
  }
  else{
    return(FALSE)
  }
}
test.tukey.outlier(tukey.outlier(x))
```

## Problem 6



## Problem 7



## Problem 8



## Problem 9



## Problem 10


## Problem 11

