---
title: "Lab 5"
author: 'Rich Chu, Lorena Mancino'
output: html_notebook
---

# Lab 5 Get Me out(lier) of Here!

## Problem 0

```{r}
x <- c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, -5.7, -7.9, -4.9, 28.7, 4.9)
```


## Problem 1

```{r}
quantile(x, 0.25)
quantile(x, 0.75)
```

## Problem 2

```{r}
quartiles <- function(x){
  first <- quantile(x, 0.25)
  third <- quantile(x, 0.75)
  list <- list(first, third, third - first)
  names(list) <- c('First', 'Third', 'Interquartile_Range')
  return(list)
}
quartiles(x)
```

## Problem 3
According to Tukey's rules, the lower bound for outliers is -18.6625 and there are no values less than that, meaning there are no lower outliers in the set.
According to Tukey's rules, the upper bound for outliers is 19.0375 and there is one value above than that, meaning that 28.7 is the only upper outliers in the set.

## Problem 4

```{r}
tukey.outlier <- function(x){
  q <- quartiles(x)
  lower.bound <- q$First - (1.5 * q$Interquartile_Range)
  upper.bound <- q$Third + (1.5 * q$Interquartile_Range)
  
  bool <- vector(length=length(x))
  
  for(i in 1:length(x)){
    if(x[i] > upper.bound){
        bool[i] <- TRUE
      }
    else if(x[i] < lower.bound){
        bool[i] <- TRUE
    }
    else{
      bool[i] <- FALSE
    }
  }
  return(bool)
}

tukey.outlier(x)
```

## Problem 5

```{r}
test.tukey.outlier <- function(FUN= tukey.outlier){
  calculated.output <- tukey.outlier(x)
  
  if(all((calculated.output == FUN) == TRUE)){
    return(TRUE)    
  }
  else{
    return(FALSE)
  }
}
test.tukey.outlier(tukey.outlier(x))
```

## Problem 6

```{r}
x2 <- -x

tukey.outlier(x2)
```
The only data value that should be an outlier in -x is the eleventh term (-28.7). This makes sense because -x is just a reflection of x, so the outlier bounds are just flipped. 

## Problem 7

```{r}

test.tukey.outlier <- function(FUN= tukey.outlier){
  calculated.output <- tukey.outlier(x)
  negative.output <- tukey.outlier(-x)
  return <- vector(length=2)
  
  if(all((calculated.output == FUN) == TRUE)){
    return[1]=TRUE    
  }
  else{
    return[1]=FALSE
  }
  
    if(all((negative.output == FUN) == TRUE)){
    return[2]=TRUE    
  }
  else{
    return[2]=FALSE
  }
  return(return)
}

```


## Problem 8

```{r}
test.tukey.outlier(tukey.outlier(x))
```


## Problem 9

```{r}
x3 <- 0.1*x

tukey.outlier(x3)
```
The only data value that should be an outlier in 0.1*x is the eleventh term (2.87). This makes sense because it's scaled down. 

## Problem 10

```{r}
test.tukey.outlier <- function(FUN= tukey.outlier){
  calculated.output <- tukey.outlier(x)
  negative.output <- tukey.outlier(-x)
  scaled.down.output <- tukey.outlier(0.1*x)
  return <- vector(length=3)
  
  if(all((calculated.output == FUN) == TRUE)){
    return[1]=TRUE    
  }
  else{
    return[1]=FALSE
  }
  
    if(all((negative.output == FUN) == TRUE)){
    return[2]=TRUE    
  }
  else{
    return[2]=FALSE
  }
  
   if(all((scaled.down.output == FUN) == TRUE)){
    return[3]=TRUE    
  }
  else{
    return[3]=FALSE
  }
  
  return(return)
}
```


## Problem 11

```{r}
test.tukey.outlier(tukey.outlier(x))
```
















